---
title: "Phylo"
author: "Heru Handika"
date: "10/15/2021"
output: html_document
---

Here, I use reshape2 to change the table format from long to wide formats and vice-versa.
Development wise, reshape2 has been superseded by pivot_wider and pivot_longer functions in the tidyr library.
But, I found reshape2 works well with matrix table that ape generated 
and produces more consistent results compare to pivot functions.

```{r}
library(here)
library(ape)
library(reshape2)
library(tidyverse)
```

## Data loading and genetic distance computing for intra-specific relationship

```{r}
data <- read.dna(here("data","cytb.phy"), format="interleaved")

distance <- dist.dna(data, model = "JC69", variance = FALSE, gamma = FALSE, pairwise.deletion = FALSE, base.freq = NULL, as.matrix = TRUE)

melted_distance <- melt(distance, value.name = "JC69")
``` 

## Computing avarage genetic distance for inter-specific relationship

```{r}
cleaned_distance <- melted_distance %>%
  separate(Var1, c("Genus1", "Species1", "Museum_no"), sep="_") %>%
  separate(Var2, c("Genus2", "Species2", "Museum_no2"), sep="_") %>%
  unite("Species", Genus1, Species1) %>%
  unite("Species2", Genus2, Species2)

# Generate inter specific relationship
inter_distance <- cleaned_distance %>% 
  group_by(Species, Species2) %>% 
  summarise(Distance = mean(JC69)) %>%
  filter(Species != Species2) # ignore intra-specific relationship

# Convert the result back to a matrix
inter_dist_matrix <- inter_distance %>% 
  dcast(Species ~ Species2, value.var = "Distance")

# Remove duplicate values in the upper triangle of the matrix
inter_dist_matrix[upper.tri(inter_dist_matrix)] <- NA

write.csv(inter_dist_matrix, here("results", "inter-specific-dist-matrix.csv"), na="")
```

